---
title: "01_find_overlap"
---
Purpose: find overlapping FoodCodes that overlap with what is available in Knight Lab's Food Tree git    
- we want to find out which foods may need to be manually curated/checked out, this mainly pertains to foods reported in ASA24-2014, which used a much older version of FNDDS.   
- we want to see what's overlapping with the Knight lab foods because there may be some foods only reported by 2014, but also exist in 2016, but just so happen to not be reported by FL100 ASA24-2016 users.     

```{r setup}
library(dplyr)
#load in Knight lab food taxonomy
mct<- read.delim("../data/00_formatted_foods/MCT_taxonomy_formatted.txt", sep="\t")
dim(mct)

#now load in the FL100 unique list of foods from ASA24-2016
items<- read.delim("../data/00_formatted_foods/ITEMS_formatted.txt", sep="\t")
dim(items)

#load the ASA24-2014 foodcodes
inf<- read.delim("../data/00_formatted_foods/INFMYPHEI_formatted.txt", sep="\t")
dim(inf)
```

clean up the food description columns
```{r}
#replace spaces with underscores
mct$Main.food.description<-trimws(mct$Main.food.description)
mct$Main.food.description<- gsub(" ", "_", mct$Main.food.description)

#remove multiple underscores
items$Food_Description<- gsub("[[:punct:]]+",'_', items$Food_Description)
inf$Food_Description<- gsub("[[:punct:]]+",'_', inf$Food_Description)

```


Remember that some of the FoodCodes are duplicated in our data because the food descriptions differ
```{r}
items.dup<-items[duplicated(items$FoodCode)|duplicated(items$FoodCode, fromLast=TRUE),]
inf.dup<-inf[duplicated(inf$FoodCode)|duplicated(inf$FoodCode, fromLast=TRUE),]

```


find overlap  
```{r}
library(gplots)
fc.venn<-venn(list("MCT" = mct$FoodID,  "Items" = items$FoodCode, "INF" = inf$FoodCode))
#887 overlap 
#141 INF + MCT only 
#1315 Items + MCT only
#101 ASA24-2014 foods do not have taxonomy assigned yet
#0 ASA24-2016 foods do not have taxonomy assigned yet 

#get the foods that need to be manually curated
inf.tmp<- data.frame(attr(fc.venn, "intersections")$'INF')
inf.man<- inf %>% filter(FoodCode %in% inf.tmp[,1])
items.tmp<-data.frame(attr(fc.venn, "intersections")$"Items")
items.man<-items %>% filter(FoodCode %in% items.tmp[,1])

venn(list("MCT" = trimws(mct$Main.food.description),  
          "Items" = trimws(items$Food_Description), 
          "INF" = trimws(inf$Food_Description)))


```

merge to get A:B:C overlap  
a = mct   
b = items  
c = inf  
```{r}
#remember some food codes are duplicated because the food descriptions differ
abc<- merge(merge(mct, items, by.x = "FoodID", by.y = "FoodCode"), inf, by.x= "FoodID", by.y="FoodCode")
length(unique(abc$FoodID)) #887
'%nin%'<-Negate('%in%')
ab<- merge(mct, items, by.x = "FoodID", by.y="FoodCode") %>% filter(FoodID %nin% inf$FoodCode)
length(unique(ab$FoodID)) #1313 correct

ac<- merge(mct,inf, by.x="FoodID", by.y="FoodCode") %>% filter(FoodID %nin% items$FoodCode)
length(unique(ac$FoodID)) #141 correct

#check
venn(list(abc$FoodID, ac$FoodID, ab$FoodID))
```

find which foods don't have matching food descriptions
```{r}
ab$check<- ifelse(as.character(ab$Main.food.description) == as.character(ab$Food_Description),"yes", "no")

ac$check<- ifelse(as.character(ac$Main.food.description) == as.character(ac$Food_Description),"yes", "no")

abc$check_ab<- ifelse(as.character(abc$Main.food.description) == as.character(abc$Food_Description.x), 1, 0)
abc$check_ac<- ifelse(as.character(abc$Main.food.description) == as.character(abc$Food_Description.y), 1, 0)
abc$check_bc<- ifelse(as.character(abc$Food_Description.x) == as.character(abc$Food_Description.y), 1, 0)
abc$check_all<- ifelse(rowSums(abc[,5:7]) == 3, 1, 0)
```

save these foods that need to be checked
```{r eval=F}
write.csv(subset(ab, check == "no"), file="../data/01_find_overlap/check_mct_items.csv", row.names=F)
write.csv(subset(ac, check=="no"), file="../data/01_find_overlap/check_mct_infmyphei.csv", row.names=F)
write.csv(subset(abc, check_all == 0), file="../data/01_find_overlap/check_mct_items_infmyphei.csv", row.names=F)
```

save the foods that need to have taxonomy manually curated   
```{r eval=F}
write.csv(inf.man, file="../data/01_find_overlap/foodcodes_manual_taxonomy_inf.csv", row.names=F)
write.csv(items.man, file="../data/01_find_overlap/foodcodes_manual_taxonomy_items.csv", row.names=F)
```

safe the duplicated foods
```{r}
write.csv(inf.dup, file="../data/01_find_overlap/duplicated_inf.csv", row.names=F)
write.csv(items.dup, file="../data/01_find_overlap/duplicated_items", row.names=F)
```

write the list with the foodcodes to help with manual curation and checking the duplicates  
```{r}
write.csv(foods2, file="../data/01_find_overlap/fl100_food_taxonomy_and_FoodCodes.csv", row.names=F )
```

