---
title: "03b_make_newick_tree_tiny"
---


Purpose: Make small newick tree of 50 foods to help MK troubleshoot opening newick tree (03/16/2020- currently not sure if the issue is that the newick file is massive or something else?)  

# prepare small list of foods (n = 50) to make the newick tree

```{r}
library(dplyr)
library(stringr)
set.seed(1234)
```

```{r}
foods<- read.delim("../data/03_make_newick_tree/fl100_final_foods.txt", sep="\t")
dim(foods) # 2413    2
```

Get n = 5-ish foods from each food group 

separate food groups
```{r}
fg1<- foods %>% filter(str_detect(FoodID, regex("^1")))
fg2<- foods %>% filter(str_detect(FoodID, regex("^2")))
fg3<- foods %>% filter(str_detect(FoodID, regex("^3")))
fg4<- foods %>% filter(str_detect(FoodID, regex("^4")))
fg5<- foods %>% filter(str_detect(FoodID, regex("^5")))
fg6<- foods %>% filter(str_detect(FoodID, regex("^6")))
fg7<- foods %>% filter(str_detect(FoodID, regex("^7")))
fg8<- foods %>% filter(str_detect(FoodID, regex("^8")))
fg9<- foods %>% filter(str_detect(FoodID, regex("^9")))
```

get a random subset of 5 or 6 foods per food group

```{r}
fg1.5 <- sample_n(fg1, 6, replace=F)
fg2.5 <- sample_n(fg2, 6, replace=F)
fg3.5 <- sample_n(fg3, 6, replace=F)
fg4.5 <- sample_n(fg4, 6, replace=F)
fg5.5 <- sample_n(fg5, 6, replace=F)
fg6.5 <- sample_n(fg6, 5, replace=F)
fg7.5 <- sample_n(fg7, 5, replace=F)
fg8.5 <- sample_n(fg8, 5, replace=F)
fg9.5 <- sample_n(fg9, 5, replace=F)

foods50<- rbind(fg1.5, fg2.5, fg3.5, fg4.5, fg5.5, fg6.5, fg7.5, fg8.5, fg9.5)

#save it
write.table(foods50, file = "../data/03_make_newick_tree/tiny_tree/fl100_tiny_foods_n50.txt", sep= "\t", quote=F)
```

# make tree


required files: 
1. nodes_fn == taxonomy, defined by FNDDS code (not per food, the general rules). Will use Knight lab file.   
2. food_database_fn == contains ASA24 FoodCodes and Food Descriptions   


```{r}
source("../lib/make.food.tree.r")
source("../food_tree_gits/Food_Tree-master/R/lib/newick.tree.r")
```

```{r}
make.food.tree(nodes="../data/03_make_newick_tree/NodeLabelsMCT.txt",
               food_database_fn = "../data/03_make_newick_tree/tiny_tree/fl100_tiny_foods_n50.txt", 
               output_tree_fn = "../data/03_make_newick_tree/tiny_tree/fl100_newick_tree_tiny_n50.txt", 
               output_taxonomy_fn = "../data/03_make_newick_tree/tiny_tree/fl100_newick_taxonomy_tiny_n50.txt")
```


