---
title: "04_get_abundance"
---

  1. Remove recalls failing the data QC done by Joanne  – this will exclude incomplete recalls but may keep subjects with kcal consumption nearly below or above the 5-95th percentiles (but still plausible)  
  2. Keep subjects consuming protein powders, cocoa powder, instant breakfast powders (i.e. the problem foods for the lactose estimation)  
  3. Keep only the subjects with at least 2 usable at-home recalls  
  4. Report abundance (grams) for each food as the average of the total n or recalls (2 or 3) per person  

import the items files 
```{r}
r14<- read.csv("../data/raw_input/INFMYPHEI_JA_Corrections.csv", header=T)
r16<- read.csv("../data/raw_input/Items_JACorrections.csv", header=T)

dim(r14)
dim(r16)
```
import the totals files, which have the data qc columns:  
```{r}
r2<- read.csv("../data/raw_input/CTSC24532USDAWHNRCNu-ASA241_DATA_2020-01-02_1128.csv", header=T)
r3<- read.csv("../data/raw_input/CTSC24532USDAWHNRCNu-ASA242_DATA_2020-01-02_1128.csv", header=T)
r4<- read.csv("../data/raw_input/CTSC24532USDAWHNRCNu-ASA243_DATA_2020-01-02_1129.csv", header=T)
```


we can't rbind the items files  because there are different ncol. select just the nutrient columns (no HEI columns) and then rbind  
```{r}
require(gplots)
require(dplyr)
#find overlapping columns 
asa<- venn(list("2014" = colnames(r14), "2016" = colnames(r16)))
cols1<- attr(asa, "intersections")$'2014:2016' 
#get the matching columns
r14.tmp<- r14[,cols1] 
r16.tmp<- r16[,cols1]
#put into one df and remove HEI columns and remove the training recall 
recalls<- rbind(r14.tmp,r16.tmp) %>% filter(RecallNo != 1)
#clean up the subject ID 
recalls$UserName<- gsub("[^0-9]", "", recalls$UserName)
```

Get the subjects passing QC 
```{r}
r2.ok<- r2 %>% filter(recallqc_passed_2 == 1) 
r3.ok<- r3 %>% filter(recallqc_passed_3 == 1) 
r4.ok<- r4 %>% filter(recallqc_passed_4 == 1)
```


Now get the items in the usable recalls  
```{r}
r2.use<- recalls %>% filter(RecallNo==2) %>% filter(UserName %in% r2.ok$subject_id)
r3.use<- recalls %>% filter(RecallNo==3) %>% filter(UserName %in% r3.ok$subject_id)
r4.use<- recalls %>% filter(RecallNo==4) %>% filter(UserName %in% r4.ok$subject_id)

#double check this is right
length(unique(r2.use$UserName)) #352
length(unique(r2.ok$subject_id)) #352
length(unique(r3.use$UserName)) #350
length(unique(r3.ok$subject_id)) #350
length(unique(r4.use$UserName)) #327
length(unique(r4.ok$subject_id)) #327
#looks good, rbind
recalls.qc<- rbind(r2.use, r3.use, r4.use)
```

Get subjects with at least 2 at home recalls 
```{r}
recalls.tmp<- unique(recalls.qc[,c("UserName", "RecallNo")])
athome<- plyr::count(recalls.tmp, "UserName") %>% filter(freq >=2) #350 subjects
#now get those subjects' items 
recalls.use<- recalls.qc %>% filter(UserName %in% athome$UserName)
#check
length(unique(recalls.use$UserName)) #350
venn(list(unique(recalls.use$UserName), unique(athome$UserName)))
venn(list("r2" = unique (r2.ok$subject_id),  #326 + 1 + 23 = 350 yay
          "r3" = unique(r3.ok$subject_id),
          "r4" = unique(r4.ok$subject_id)))
```

`recalls.use` contains all the items for the subjects we can use for analysis   
we now need to average the g of food over the total n of recalls per subject   

some examples to check:   

5001 ate 100.5 of FoodCode 22311000 in Recall 2  
5001 has 3 at home recalls, so the average of 22311000 is 100.5/3 = 33.5  

6018 consumed 11100000 twice (recalls 3 and 4), each at 61 g for a grand total of 122
6018 has 3 at home recalls, so the average is 122/3 = 40.67


```{r}
# sum total grams of each food eaten within a record
otu.total <- aggregate(FoodAmt ~ UserName +FoodCode, data=recalls.use,FUN="sum")
#now we have to merge it with the frequency or else aggregate won't divide by the total n of recalls (since not all foods appear in every recall) 
otu.freq<- merge(otu.total, athome, by = "UserName")
otu.freq$avg <- otu.freq$FoodAmt/otu.freq$freq
```

reshape 
```{r}
otu <- otu.freq %>% select(UserName, FoodCode, avg)
otu.w <- reshape(otu, timevar = "FoodCode", idvar = "UserName", direction = "wide")
otu.w[is.na(otu.w)] <- 0
rownames(otu.w) <- otu.w[,1] # make subjectID the rownames
otu.w <- otu.w[,-1]    
colnames(otu.w) <- gsub("avg.", "", colnames(otu.w)) #rename column names to FoodCodes only
t.otu.w <- t(otu.w)#transpose
```

