---
title: "04c_average_recalls"
---

purpose: to average the food amounts over 2 or 3 recalls   

Previous QC'ing of recalls included:  
  * Recalls were excluded if they failed QC  
  * Subjects were excluded if they had fewer than 2 at-home recalls   

some examples to check:   

5001 ate 100.5 of FoodCode 22311000 in Recall 2  
5001 has 3 at home recalls, so the average of 22311000 is 100.5/3 = 33.5  

6018 consumed 11100000 twice (recalls 3 and 4), each at 61 g for a grand total of 122
6018 has 3 at home recalls, so the average is 122/3 = 40.67


```{r}
#import data 
recalls.use<- read.csv("../data/04_get_abundance/recalls_foodcodes_qcd.csv", header=T) #19373 x 111, correct n of rows. 
```

```{r}
# sum total grams of each food eaten within a record
otu.total <- aggregate(FoodAmt ~ UserName +FoodCode, data=recalls.use,FUN="sum")
#now we have to merge it with the frequency or else aggregate won't divide by the total n of recalls (since not all foods appear in every recall) 
recalls.tmp<- unique(recalls.use[,c("UserName", "RecallNo")])
athome<- plyr::count(recalls.tmp, "UserName") #350 subjects
otu.freq<- merge(otu.total, athome, by = "UserName")
otu.freq$avg <- otu.freq$FoodAmt/otu.freq$freq
```

```{r}
otu.save<- otu.freq %>% select(UserName, FoodCode, avg)
colnames(otu.save)<-c("UserName", "FoodCode", "FoodAmt") #the 'avg' column needs to be named "FoodAmt" for the below function
write.table(otu.save, "../data/04_get_abundance/recalls_average_g_041420.txt", sep="\t", row.names=F, quote=F)
```

```{r}
#sanity check-- how many unique foods are in here? 
length(unique(recalls.use$FoodCode)) #2129
length(unique(otu.save$FoodCode)) #2125

'%nin%'<- Negate('%in%')
which<- recalls.use %>% filter(FoodCode %nin% unique(otu.save$FoodCode)) %>% select(FoodCode, Food_Description) %>% unique()
#correct, the 4 foodcodes are generated from ASA24 system errors and all had NA values for nutrients etc. hooooray
```


```{r}
source("../lib/make.food.otu.R")

make.food.otu(food_records_fn = "../data/04_get_abundance/recalls_average_g.txt" ,
              food_record_id = "UserName", 
              food_taxonomy_fn= "../data/03_make_newick_tree_all_foods/fl100_newick_taxonomy_all_foods.txt", 
              output_fn="../data/04_get_abundance/fl100_otu_abundance_041420.txt")

```

