---
title: "05_make_newick_tree_recall_foods"
---


Get only the foodcodes/food descriptions for the foods reported by subjects passing recall qc (2 or 3 usable recalls):  

We can do this by using the output of the recall QC step, which contains the SubjectID, FoodCode, and average FoodAmt but not food description (recalls_average_g.txt). We can import the final list of **all** FL100 FoodCodes, which contains the FoodCodes and Food Description (fl100_final_foods.txt), and is already formatted for input into the make.food.tree function.  
We find the overlap of the two files, using the file structure of the FL100 FoodCodes file. This will be used as input for a new tree. 


```{r}
#input for otu table-- from QC'd recalls
amts<- read.delim("../data/04_get_abundance/recalls_average_g.txt", header=T, sep = "\t") #13339 x 3 correct
colnames(amts)
length(unique(amts$UserName)) #350, correct
length(unique(amts$FoodCode)) #2125, correct 

```

```{r}
#input of the list of all QC'd fl100 foods and descriptions 
foodlist<- read.delim("../data/03_make_newick_tree_all_foods/fl100_final_all_foods.txt", sep = "\t", header=T) #2412 x 2 correct
colnames(foodlist)
length(unique(foodlist$FoodID)) #2412 correct
length(unique(foodlist$Main.food.description)) #2412
```

now find the overlap 
```{r}
overlap<- foodlist %>% filter(FoodID %in% amts$FoodCode)
dim(overlap) #2125, correct-- it's the same n rows as 'amts' 

#double check
venn(list(amts$FoodCode, overlap$FoodID)) #yep
venn(list(unique(amts$FoodCode), unique(overlap$FoodID))) #yep
venn(list(unique(amts$FoodCode), unique(foodlist$FoodID))) #2125 overlap, 287 foods are not included in qc'd recalls 

#save this 
write.table(overlap, "../data/05_make_newick_tree_foods_from_recalls/fl100_final_foods_from_recalls.txt", sep="\t", quote = F, row.names=F)
```


required files: 
1. nodes_fn == taxonomy, defined by FNDDS code (not per food, the general rules). Will use Knight lab file.   
2. food_database_fn == contains ASA24 FoodCodes and Food Descriptions   

# make tree

```{r}
source("../lib/make.food.tree.r")
source("../food_tree_gits/Food_Tree-master/R/lib/newick.tree.r")
```

```{r}
make.food.tree(nodes="../data/03_make_newick_tree_all_foods/NodeLabelsMCT.txt",
               food_database_fn = "../data/05_make_newick_tree_foods_from_recalls/fl100_final_foods_from_recalls.txt", 
               output_tree_fn = "../data/05_make_newick_tree_foods_from_recalls/fl100_newick_tree_from_recalls.txt", 
               output_taxonomy_fn = "../data/05_make_newick_tree_foods_from_recalls/fl100_newick_taxonomy_from_recalls.txt")
```


for a sanity check, re-make an otu table using this exported taxonomy list and compare the outputs to make sure they are identical

```{r}
source("../lib/make.food.otu.R")

make.food.otu(food_records_fn = "../data/04_get_abundance/recalls_average_g.txt" ,
              food_record_id = "UserName", 
              food_taxonomy_fn= "../data/05_make_newick_tree_foods_from_recalls/fl100_newick_taxonomy_from_recalls.txt", 
              output_fn="../data/05_make_newick_tree_foods_from_recalls/fl100_otu_abundance_tmp_recallsonly.txt")

```

```{r}
og_otu<- read.delim("../data/04_get_abundance/fl100_otu_abundance.txt", sep = "\t", header=T)
new_otu<- read.delim("../data/05_make_newick_tree_foods_from_recalls/fl100_otu_abundance_tmp_recallsonly.txt", sep = "\t", header=T)

identical(og_otu, new_otu) #TRUE we're good 
```

is it identical to the file sent to mary 02/11/2020?
```{r}
mary<- read.delim("../data/04_get_abundance/fl100_otu_abundance_sent_to_mary_021120.txt", sep= "\t", header=T)

identical(mary, new_otu) #FALSE
identical(mary[,2:351], new_otu[,2:351]) #values are identical

diffs<- anti_join(mary[,c(1,352)], new_otu[,c(1,352)])
#the difference is because: in 02b_cleanup_foodcodes.Rmd, lines 100-103 selects just one example of each foodcode (after QC),for some reason this time it picked different versions of: 
# new: Vegetable beef soup canned prepared with water or readytoserve
# old: Vegetable beef soup prepared with water 
# new: Soft drink fruit flavored caffeine free
# old: Soft drink fruitflavored caffeine free	
```
```{r}
another_otu<- read.delim("../data/04_get_abundance/fl100_otu_abundance_041420.txt", sep = "\t", header = T)
identical(another_otu, new_otu)
identical(another_otu[,2:351], mary[,2:351])
```

