---
title: "07_newicktree_nowater"
---

Make a newick tree with no water (L2: Water_noncarbonated )

After reviewing the input otu table, I see that 9400000 (Water used as an ingredient) and 20000000 (Meat, NFS) do not have L2-4 because their foodcodes get put into scientific notation and even though I change foodcodes with scientific = F, they still don't get recognized. 94000000 will get removed because it should be part of L2 water_noncarbonated. I will manually place Meat,NFS to L2: Meat, NS_as_to_type in the food tree 

```{r}
set.seed(0)
library(dplyr)
library(gplots)
#input for otu table-- from QC'd recalls
amts<- read.delim("../data/04_get_abundance/recalls_average_g.txt", header=T, sep = "\t") #13339 x 3 correct
colnames(amts)
length(unique(amts$UserName)) #350, correct
length(unique(amts$FoodCode)) #2125, correct 

#fix the food codes coming up as scientific notation 
amts$FoodCode_tmp<- as.integer(format(amts$FoodCode, scientific = F))
for(i in 1:nrow(amts)){
  if(amts[i,"FoodCode_tmp"] == "94000000"){
    amts[i,"FoodCode"] <- as.integer(94000000)
  }else if(amts[i,"FoodCode_tmp"] == "20000000"){
    amts[i,"FoodCode"] <- as.integer(20000000)
  }
}

#The food description is missing 
```

remove subjects consuming water variabiles (^94 and ^Water)
```{r}
library(stringr)
water<- amts %>% filter(str_detect(as.integer(FoodCode), regex("^94"))) #415 subjects
nowater<- amts %>% filter(!str_detect(as.integer(FoodCode), regex("^94"))) #12924 subjects, correct, 2122 unique foodcodes
write.table(nowater, "../data/04_get_abundance/recalls_average_g_nowater.txt",  sep="\t", row.names=F, quote=F)
```

get the taxonomy
```{r}
tax<- read.delim("../data/05_make_newick_tree_foods_from_recalls/fl100_newick_taxonomy_from_recalls.txt", sep = "\t", header=T, colClasses = "character")

tax_nowater<- tax %>% filter(!str_detect(as.integer(FoodID), "^94"))

write.table(tax_nowater, "../data/05_newicktree_nowater/fl100_taxonomy_nowater_for_otu.txt", sep="\t", quote = F, row.names=F)
```

```{r}
source("../lib/make.food.otu.R")

make.food.otu(food_records_fn = "../data/04_get_abundance/recalls_average_g.txt" ,
              food_record_id = "UserName", 
              food_taxonomy_fn= "../data/05_newicktree_nowater/fl100_taxonomy_nowater_for_otu.txt", 
              output_fn="../data/05_newicktree_nowater/fl100_otu_abundance_nowater.txt")

```




# food list
```{r}
foodlist<- read.delim("../data/05_make_newick_tree_foods_from_recalls/fl100_final_foods_from_recalls.txt", sep = "\t", header=T, 
                      colClasses = c("character", "character")) #2125 x 2 correct (2125 unique rows)
#remove water
foodlist_nowater<- foodlist %>% filter(!str_detect(as.integer(FoodID), regex("^94"))) #meat NFS is still messed up

for(i in 1:nrow(foodlist_nowater)){
  if(foodlist_nowater[i, "Main.food.description"] == "Meat_NFS"){
    foodlist_nowater[i,"FoodID"] <- as.integer(20000000)
  }
}


write.table(foodlist_nowater, file = "../data/05_newicktree_nowater/fl100_foods_nowater.txt", sep="\t", quote = F, row.names=F)
```

required files: 
1. nodes_fn == taxonomy, defined by FNDDS code (not per food, the general rules). Will use Knight lab file.   
2. food_database_fn == contains ASA24 FoodCodes and Food Descriptions   

# make tree

```{r}
source("../lib/make.food.tree.r")
source("../food_tree_gits/Food_Tree-master/R/lib/newick.tree.r")
```

```{r}
make.food.tree(nodes="../data/03_make_newick_tree_all_foods/NodeLabelsMCT.txt",
               food_database_fn = "../data/05_newicktree_nowater/fl100_foods_nowater.txt", 
               output_tree_fn = "../data/05_newicktree_nowater/fl100_newick_tree_nowater.txt", 
               output_taxonomy_fn = "../data/05_newicktree_nowater/fl100_newick_taxonomy_nowater.txt")
```


