---
title: "08_NewickTree_w_FilteredSamples"
---

This notebook builds a food tree/taxonomy/otu table (w/ water removed) after filtering to remove the following subjects, as requested by MK 06/12/2020:   

5022
6066
6072
6082
6091
7063
8012


```{r}
library(dplyr)
avgs<- read.delim("../data/04_get_abundance/recalls_average_g_nowater.txt", header=T, sep = "\t") #this has water removed
#12924 x 4
length(unique(avgs$UserName)) #350 unique subjects 
length(unique(avgs$FoodCode)) #2122 unique foods 

#fix the food codes coming up as scientific notation 
avgs$FoodCode_tmp<- as.integer(format(avgs$FoodCode, scientific = F))
for(i in 1:nrow(avgs)){
  if(avgs[i,"FoodCode_tmp"] == "94000000"){
    avgs[i,"FoodCode"] <- as.integer(94000000)
  }else if(avgs[i,"FoodCode_tmp"] == "20000000"){
    avgs[i,"FoodCode"] <- as.integer(20000000)
  }
}

subjrm<- read.delim("../data/06_filter_samples/FoodSamplesToDiscard.txt", header=F, sep = "\t") #these are the subjects to remove
subjrm <- as.character(subjrm$V1)
subjrm #7 subjects 
```

```{r}
'%nin%' <- Negate("%in%")
#filter the averages table to remove subjects 
filt<- avgs %>% filter(UserName %nin% subjrm)
length(unique(filt$UserName)) #343, correct- 7 subjects removed 
length(unique(filt$FoodCode)) #2107 unique foods 

write.table(filt, "../data/06_filter_samples/recalls_average_g_filt.txt",  sep="\t", row.names=F, quote=F)
```

```{r}
source("../lib/make.food.otu.R")

make.food.otu(food_records_fn = "../data/06_filter_samples/recalls_average_g_filt.txt" ,
              food_record_id = "UserName", 
              food_taxonomy_fn= "../data/05_newicktree_nowater/fl100_taxonomy_nowater_for_otu.txt", 
              output_fn="../data/06_filter_samples/fl100_otu_abundance_filt.txt")
```

check the output of the otu table you just made 
```{r}
otu<- read.delim("../data/06_filter_samples/fl100_otu_abundance_filt.txt", header=T, sep = "\t", stringsAsFactors = F)
dim(otu) #2107 foods x 345 cols (343 subjects + Taxconomny & FoodCode), correct
otu %>% dplyr::select(X.FoodCode, taxonomy) %>% View() #Meat, NFS taxonomy is incorrect. 
```

Fix Meat, NFS taxonomy in OTU table: 

currently: 
L1_Meat_Poultry_Fish_and_Mixtures;L2_;L3_;L4_;L5_;Meat_NFS

Correct: 
L1_Meat_Poultry_Fish_and_Mixtures;L2_Meat;L3_Meat_NS_as_to_type;L4_;L5_;Meat_NFS

```{r}
tmp<- otu
for(i in 1:nrow(tmp)){
  if(tmp[i, "X.FoodCode"] == 'Meat NFS'){
    tmp[i, "taxonomy"] <- "L1_Meat_Poultry_Fish_and_Mixtures;L2_Meat;L3_Meat_NS_as_to_type;L4_;L5_;Meat_NFS"
  }
}

tmp %>% dplyr::select(X.FoodCode, taxonomy) %>% filter(X.FoodCode == "Meat NFS") %>% View() #looks good

write.table(tmp, "../data/06_filter_samples/fl100_otu_abundance_filt.txt", row.names = F, sep = "\t", quote=F)
```


# get a unique list of foods 

```{r}
#import the food list w/ water fixed. this contains all the unique foods and food descriptions 
foodlist<- read.delim("../data/05_newicktree_nowater/fl100_foods_nowater.txt", header=T, sep = "\t") #2122 foods, same as above 

#get a unique list of foodcodes from the list of filtered subjects 
foodfilt<- unique(filt$FoodCode) #2107 

#now keep only the foods in the filtered food list
foodkeep<- foodlist %>% filter(FoodID %in% foodfilt)
length(unique(foodkeep$FoodID)) #2107 correct

for(i in 1:nrow(foodkeep)){
  if(foodkeep[i, "Main.food.description"] == "Meat_NFS"){
    foodkeep[i,"FoodID"] <- as.integer(20000000)
  }
}


write.table(foodkeep, "../data/06_filter_samples/fl100_foods_filt.txt",sep="\t", quote = F, row.names=F)
```

required files: 
1. nodes_fn == taxonomy, defined by FNDDS code (not per food, the general rules). Will use Knight lab file.   
2. food_database_fn == contains ASA24 FoodCodes and Food Descriptions   

# make tree

```{r}
source("../lib/make.food.tree.r")
source("../food_tree_gits/Food_Tree-master/R/lib/newick.tree.r")
```

```{r}
make.food.tree(nodes="../data/03_make_newick_tree_all_foods/NodeLabelsMCT.txt",
               food_database_fn = "../data/06_filter_samples/fl100_foods_filt.txt", 
               output_tree_fn = "../data/06_filter_samples/fl100_newick_tree_filt.txt", 
               output_taxonomy_fn = "../data/06_filter_samples/fl100_newick_taxonomy_filt.txt")
```
_
check the outputs 
```{r}
tax_filt<- read.table("../data/06_filter_samples/fl100_newick_taxonomy_filt_tmp.txt", header=T, sep = "\t", stringsAsFactors = F)
dim(tax_filt) #2107 x 3
#Meat, NFS taxonomy is correct and needs to be manually curated. 
```

